Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-08-05 Paper/assa.tex
...

\subsection{Active subspaces}
\label{sub:ac}
As mentioned in the introduction, the active subspace is low-dimensional subspace
that constitutes important directions in a model's input
space~\cite{Constantine:2015}. The effective variability in the model output, $f$
due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
along these directions. Each component of $\bm{\theta}$
is mapped to a corresponding value in $\Omega\in\mathbb{R}^{N_p}$
($N_p$: number of uncertain inputs)
using a linear transformation.  The set of canonical random variables, corresponding
to $\theta_i$'s, in $\Omega$ is referred to as $\bm{\xi}$.
Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
these directions are the eigenvectors of a positive
semi-definite matrix, $\mat{C}$, defined as follows: 
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where 
$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
Here 
$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
\ee
%
where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
\] 
The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2 \\
\end{bmatrix}
\ee
 %
 where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
eigenvalues. Once the active subspace is computed, dimension reduction is
accomplished by transforming the set of canonical random variables, $\bm{\xi}$
into a reduced set, $\mat{W}_1^\top \bm{\xi}$, referred to as the set of active
variables. The number of active variables is thus equal to the number
of dominant eigenvectors. The model output in the original space, $f(\bm{\xi}(\theta))$,
is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
In practice, the matrix, $\mat{C}$ is typically
approximated using pseudo-random sampling techniques such as Monte Carlo and
Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
 = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
\label{eq:chat}
 \ee
 %
Clearly, the amount of computational effort associated with constructing the matrix,
$\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
computational approach is adopted in this work to ensure that $N$ is sufficiently
large so that the dominant eigenvectors in $\hat{\mat{W}}_1$ are approximated
with reasonable accuracy as discussed later in Section~\ref{sec:method}.

>>>>

\subsection{Active subspaces}
\label{sub:ac}
As mentioned in the introduction, the active subspace is low-dimensional subspace
that constitutes important directions in a model's input
space~\cite{Constantine:2015}. The effective variability in the model output, $f$
due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
along these directions. Each component of $\bm{\theta}$
is mapped to a corresponding value in $\Omega\in\mathbb{R}^{N_p}$
($N_p$: number of uncertain inputs)
using a linear transformation.  The set of canonical random variables, corresponding
to $\theta_i$'s, in $\Omega$ is referred to as $\bm{\xi}$.
Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
these directions are the eigenvectors of a positive
semi-definite matrix, $\mat{C}$, defined as follows: 
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where 
$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
Here 
$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
\ee
%
where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
\] 
The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2 \\
\end{bmatrix}
\ee
 %
 where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
eigenvalues. Once the active subspace is computed, dimension reduction is
accomplished by transforming the set of canonical random variables, $\bm{\xi}$
into a reduced set, $\mat{W}_1^\top \bm{\xi}$, referred to as the set of active
variables. The number of active variables is thus equal to the number
of dominant eigenvectors. The model output in the original space, $f(\bm{\xi}(\theta))$,
is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
\alennote{This is too fast, explain a bit more, and refer to 
Paul Constantine's book. I think we should follow the explanations
in Chapter 4 of Paul's book; see equations (4.1)--(4.3) and algorithm 4.1
for surrogate model construction on active subspace; these should be
mentioned here.}
In practice, the matrix, $\mat{C}$ is typically
approximated using pseudo-random sampling techniques such as Monte Carlo and
Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
 = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
\label{eq:chat}
 \ee
 %
Clearly, the amount of computational effort associated with constructing the matrix,
$\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
computational approach is adopted in this work to ensure that $N$ is sufficiently
large so that the dominant eigenvectors in $\hat{\mat{W}}_1$ are approximated
with reasonable accuracy as discussed later in Section~\ref{sec:method}.

### 2018-08-05 Paper/method.tex
...

In this section, we present the computational strategy used for computing the active subspace in the case of
H$_2$/O$_2$ reaction kinetics with uncertain pre-exponent of individual reaction rates provided in
Table~\ref{tab:kinetics}. Specifically, we employ two strategies: One that involves computation of model
gradients in order to construct the matrix (gradient-based), $\hat{\mat{C}}$ in~\eqref{eq:chat}, and the other that involves a local 
linear approximation of the model output to circumvent the computational cost associated with gradient
computation (gradient-free). The gradient-based approach is based on Algorithm 1.1, and the gradient-free approach is
based on Algorithm 1.2 in~\cite{Constantine:2015}. However, in this work, we have implemented the
two strategies in an iterative manner and have demonstrated potential for additional savings in both cases.
Additionally, for the present application, our findings based on the two strategies are observed to be consistent.
Since, the gradient-free approach is relatively more efficient, this observation encourages its use for high-dimensional
applications as discussed later in Section~\ref{sec:app}. Note however, that the model output is required to be
differentiable throughout the input domain in both cases.

@@

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$, at an initial set of $n_1$ samples. 
For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
estimate of the dominant eigenspace, $\hat{\mat{W}}_1$ and corresponding set of eigenvalues, $\hat{\mat{\Lambda}}_1$.
At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
of gradient evaluations are augmented with the available set to re-construct $\hat{\mat{C}}$ followed by its eigenvalue
decomposition. The relative
change in the norm of the difference in squared value of individual components of the dominant eigenvectors between subsequent 
iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$, is smaller a given tolerance, $\tau$. 
A regression fit to
 $G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
 output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores,
 $\bm{\nu}_r(f)$, which provide an insight into the relative importance of the uncertain inputs. Note that the
 index, $r$, corresponds to the number of dominant eigenvectors in $\hat{\mat{W}}_1$. The sequence of 
 steps for the gradient-based approach are provided in Algorithm~\ref{alg:grad}.

@@

In order to motivate the iterative strategy used in this work, we implement Algorithm~\ref{alg:grad} to the
19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
verification, $\hat{\mat{C}}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
the comparison of the resulting normalized eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
relative L-2 norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
Interestingly, we observe that the dominant eigenvalues ($\lambda_1 - \lambda_4$) are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvectors. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we explore the applicability of a so-called gradient-free approach as mentioned earlier, in the
following section.

@@

As mentioned earlier, the gradient-free approach does not involve computation of the gradient of the model
output. Instead, a least-squares fit is performed to the local linear regression-based approximation
to the model output. While the gradient-free approach offers a significant potential for additional 
computational savings, its ability to reasonably approximate the active subspace 
has not been investigated for a wide variety of problems~\cite{Constantine:2015}. Nevertheless, we are motivated by the
underlying potential of this approach in reducing the computational effort associated with the
active subspace computation. Moreover, this effort could help advance this approach and further motivate
its usage for a wide variety of high-dimensional applications in the future.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
$f(\bm{\xi})$ can be transformed as a function of the active variable as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
compared for the two approaches in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the gradient-based and the gradient-free approaches.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the variability in ignition
delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
active subspace; consistent with the fact the it is 1-dimensional.

We extend the comparative assessment of the two approaches by estimating the activity score for individual
uncertain parameters ($\alpha_i$) using
the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}. The activity scores for the 19 
uncertain pre-exponents ($A_i$'s), estimated using
the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 
compare these estimates with those obtained using the DGSMs~\cite{Vohra:2018} in 
Figure~\ref{fig:comp_as} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the
Poincare constant and the DGSM ($\mu_i$) (adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores estimated using the two approaches agree favourably with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons establish consistency between the two approaches. Note that the computational
effort in the case of gradient-free approach is significantly smaller. Hence, we are encouraged to use the
gradient-free approach for the 33-dimensional application in the next section.

>>>>

In this section, we present the computational strategy used for computing the
active subspace in the case of H$_2$/O$_2$ reaction kinetics with uncertain
pre-exponent of individual reaction rates provided in Table~\ref{tab:kinetics}.
Specifically, we employ two strategies: 
A gradient-based strategy that involves computation of model
gradients in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a gradient-free approach 
that involves a (\alennote{should we say 
regression-based?}) local linear approximation of
the model output to circumvent the computational cost associated with gradient
computation. The gradient-based approach is based on Algorithm
1.1, and the gradient-free approach is based on Algorithm 1.2
in~\cite{Constantine:2015}. In this work, we implement the two
strategies in an iterative manner and demonstrate potential for
additional savings in both cases.  For the present application,
our findings based on the two strategies are observed to be consistent.  
Since,
the gradient-free approach is more efficient, this observation
encourages its use for high-dimensional applications as discussed later in
Section~\ref{sec:app}. 
%Note however, that the model output is required to be
%differentiable throughout the input domain in both cases.

@@

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
at an initial set of $n_1$ samples.  For this purpose, $\bm{\xi}_i$'s are
projected to the physical space as $\bm{\theta}_i$'s. 
\alennote{The mention of $\bm{\theta}$ is redundant. 
Say what $\bm{\xi}_i$, $i = 1, \ldots, n_1$ are before using them.} 
Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
$\hat{\mat{W}}_1$ and corresponding set of eigenvalues,
$\hat{\mat{\Lambda}}_1$.  At each subsequent iteration, model evaluations are
generated at a new set of $n_s$ samples. The new set of gradient evaluations
are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
by its eigenvalue decomposition. The relative change in the norm of the
difference in squared value of individual components of the dominant
eigenvectors between subsequent iterations is evaluated. The process is
terminated and the resulting eigenspace is considered to be converged once
highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$,
is smaller than a given tolerance, $\tau$.  A regression fit to
$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
quantify the uncertainty in the model output. Moreover, the components of the
dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
which provide insight into the relative importance of the uncertain inputs.
Note that the index, $r$, corresponds to the number of dominant eigenvectors in
$\hat{\mat{W}}_1$. The sequence of steps for the gradient-based approach are
provided in Algorithm~\ref{alg:grad}.

@@

To motivate our iterative strategy, we implement
Algorithm~\ref{alg:grad} to the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain $A_i$'s. For the purpose of verification,
$\hat{\mat{C}}$ is initially constructed using a large set of samples
($N$~=~1000) in the input domain and the corresponding active subspace is
computed. In Figure~\ref{fig:eig_comp}~(left), we illustrate the comparison of
the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
\lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$.
Additionally, in Figure~\ref{fig:eig_comp} (right), we compare relative $L_2$
norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
\alennote{This last sentence is unclear.}

%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative $L_2$ 
norm of the difference between
the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvectors. In this case, since 
$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we next explore the applicability of a so-called gradient-free approach
as mentioned earlier.

@@

\alennote{I don't see the point of this paragraph; part of it that 
explains the gradient-free approach is unclear, and the part 
that seems to motivate its use is something to mention in the introduction.
Just say what the gradient-free approach is about, and say that we 
follow an iterative strategies as before.}
As mentioned earlier, the gradient-free approach does not involve computation
of the gradient of the model output. Instead, a least-squares fit is performed
to the local linear regression-based approximation to the model output. While
the gradient-free approach offers a significant potential for additional
computational savings, its ability to reasonably approximate the active
subspace has not been investigated for a wide variety of
problems~\cite{Constantine:2015}. We are motivated by the
potential of this approach in reducing the computational effort
associated with the active subspace computation. Moreover, this effort could
help advance this approach and further motivate its usage for a wide variety of
high-dimensional applications in the future.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
$f(\bm{\xi})$ can be transformed as a function of the active variable as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
compared for the two approaches in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the gradient-based and the gradient-free approaches.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the variability in ignition
delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
active subspace; consistent with the fact the it is 1-dimensional.
\alennote{The last sentence is incomplete; also clarify that ``it'' refers to 
the active subspace.}

We complement the comparative assessment of the two approaches by estimating
the activity scores for individual uncertain parameters ($\alpha_i$) using the
components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores (\alennote{relative activity scores?}) for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the gradient-based and gradient-free approaches are plotted in
Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
verification, we compare these estimates with those obtained using the
corresponding DGSMs, as reported in~\cite{Vohra:2018}, in
Figure~\ref{fig:comp_as} (right).

%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the
Poincare constant and the DGSM ($\mu_i$) (adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores estimated using the two approaches agree favourably with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons establish consistency between gradient-based and
gradient-free approaches. Due to its reduced computational overhead, the
gradient-free approach is our method of choice for the 33-dimensional problem
in the next section.
